---
title: "Model selection"
author: "Dr. Çetinkaya-Rundel"
date: "October 19, 2017"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/sta112-hex.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, show.signif.stars = FALSE)
blue = "#425460"
library(emo)
```

class: center, middle

# Getting started

---

## Getting started

- Any questions from last time?

- Quality of fit and model selection

---

class: center, middle

# Quality of fit in MLR

---

## Data: Paris Paintings

From last time...

```{r message=FALSE, warning = FALSE}
library(tidyverse) # ggplot2 + dplyr + readr + and some others
library(broom)
```

```{r message=FALSE}
# Load data
pp <- read_csv("data/paris_paintings.csv", na = c("n/a", "", "NA"))

# Filter for paintings less than 5000 square inches
pp_Surf_lt_5000 <- pp %>%
  filter(Surface < 5000)
  
# Fit model with main effects only
m_main <- lm(log(price) ~ Surface + factor(artistliving), 
             data = pp_Surf_lt_5000)
             
# Fit model with main effects and interactions
m_int <- lm(log(price) ~ Surface + factor(artistliving) +
                         Surface * factor(artistliving), 
            data = pp_Surf_lt_5000)
```

---

## $R^2$

- $R^2$ is the percentage of variability in the response variable explained by the 
regression model.

```{r}
glance(m_main)$r.squared
glance(m_int)$r.squared
```

--

- Clearly the model with interactions has a higher $R^2$.

--

- However using $R^2$ for model selection in models with multiple explanatory 
variables is not a good idea as $R^2$ increases when **any** variable is added to the 
model.

---

## $R^2$ - first principles

$$ R^2 =  \frac{ SS\_{Reg} }{ SS\_{Total} } = 1 - \left( \frac{ SS\_{Error} }{ SS\_{Total} } \right) $$

<div class="question">
Calculate $R^2$ based on the output below.
</div>

```{r digits = 3}
anova(m_main)
```

---

## Adjusted $R^2$

$$ R^2\_{adj} = 1 - \left( \frac{ SS\_{Error} }{ SS\_{Total} } \times \frac{n - 1}{n - k - 1} \right), $$

where $n$ is the number of cases and $k$ is the number of predictors in the model

--

- Adjusted $R^2$ doesn't increase if the new variable does not provide any new 
informaton or is completely unrelated.

--

- This makes adjusted $R^2$ a preferable metric for model selection in multiple
regression models.

---

## In pursuit of Occam's Razor

- Occam's Razor states that among competing hypotheses that predict equally well, 
the one with the fewest assumptions should be selected.

- Model selection follows this principle.

- We only want to add another variable to the model if the addition of that
variable brings something valuable in terms of predictive power to the model.

- In other words, we prefer the simplest best model, i.e. **parsimonious** model.

---

## Comparing models

It appears that adding the interaction actually increased adjusted $R^2$, so we 
should indeed use the model with the interactions.

```{r}
glance(m_main)$adj.r.squared
glance(m_int)$adj.r.squared
```

---

class: center, middle

# Model selection

---

## Backwards elimination

- Start with **full** model (including all candidate explanatory variables and all
candidate interactions)

- Remove one variable at a time, and select the model with the highest adjusted $R^2$

- Continue until adjusted $R^2$ does not increase

---

## Forward selection

- Start with **empty** model

- Add one variable (or interaction effect) at a time, and select the model with the 
highest adjusted $R^2$

- Continue until adjusted $R^2$ does not increase

---

## Model selection and interaction effects

If an interaction is included in the model, the main effects of both of
those variables must also be in the model

---

## Other model selection criteria

- Adjusted $R^2$ is one model selection criterion

- There are others out there (many many others!), we'll discuss some later in the 
course, and some you might see in another courses

