---
title: "Case study 04 - Inferring from the GSS, Pt 1"
date: "Due at the end of class on Nov 14"
author: "Team assignment"
output: 
  html_document: 
    theme: yeti
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Accessing the data

In this case study, and in the subsequent mini homework, you will work with the 2016
General Social Survey (GSS). The data can be found in the `\data` folder of your 
assignment repository. This is an excerpt from the 2016 GSS containing only the variables 
that will be used for these two assignments. We're not distributing the entire dataset
in order to keep the size of the dataset reasonable.

You can load the data with

```{r eval=FALSE}
gss16 <- read_csv("data/gss16_excerpt.csv")
```

Some of the questions ask about conditions for inference. Note that the GSS employs 
random sampling.

### Accessing the assignment repo

Go to the #assignment-links channel on Slack, click on the link, and accept the assignment.

### Task

First focus is Americans' "belief" in evolution. The GSS asks

> Human beings, as we know them today, developed from earlier species of animals. (Is that true or false?)

The responses to this question are stored in the variable called `EVOLVED`.

1. Create a new data frame that omits `NA` values for the `EVOLVED` variable. How many observations are in this data frame?
2. Summarise the distribution of responses to this question.
3. Is the independence condition satisfied for these data? Explain your reasoning.
4. Estimate and interpret the proportion of Americans who believe in evolution with a 95% confidence interval...
    a. Using simulation based methods.
    b. Using CLT based methods (only if the sample size condition is satisfied). You will need to use the `prop.test` function to build this interval.

    ```{r eval=FALSE}
    prop.test(x = [number of successes], n = [sample size], conf.level = 0.95)
    ```

5. Do your intervals produce roughly similar results?
6. If we were to use 99% confidence level, would the widths of your intervals be the same, larger, or smaller as the widths of the respective 95% confidence intervals?
7. Construct the 99% confidence intervals, and comment on whether your answer from part (e) is confirmed.

### Grading

Check / no check