---
title: "Mini HW 09 - Getting started with modeling Paris Paintings"
date: "Due: Oct 17 at 10:05am"
author: "Team assignment"
output: 
  html_document: 
    theme: yeti
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Accessing the data

The data can be found in the `\data` folder of your homework repository.

Read the data with

```{r eval=FALSE}
pp <- read_csv("data/paris_paintings.csv")
```

Take a peek at the [codebook](https://www2.stat.duke.edu/courses/Fall15/sta112.01/data/paris_paintings.html).

### Accessing the assignment repo

Go to the #assignment-links channel on Slack and click on the link for mini-hw-07, and accept the assignment. Note that this is an individual assignment.

### Assignment

The models you fit should not be the exact same ones we reviewed in class (but they can share *some* explanatory variables, just not all).

1. Fit a linear model for predicting log(price) of paintings based on a single numerical variable.

- Find the slope and intercept of the model, and interpret each in context of the data.
- Find and interpret the $R^2$ of the model.

2. Fit a linear model for predicting log(price) of paintings based on a single categorical variable with two levels.

- Find the slope and intercept of the model, and interpret each in context of the data.
- Find and interpret the $R^2$ of the model.

3. Fit a linear model for predicting log(price) of paintings based on a single categorical variable with more than two levels. If there are too many levels, consider collapsing down to a reasonable number first, then fit the model.

- Find the slopes and intercept of the model, and interpret each in context of the data.
- Find and interpret the $R^2$ of the model.

4. Fit a linear model for predicting log(price) of paintings based on at least two variables (and not a ridiculous amount, be selective!). If some of your categorical variables have too many levels, consider collapsing down to a reasonable number first, then fit the model.

- Find the slopes and intercept of the model, and interpret each in context of the data.
- Find and interpret the $R^2$ of the model.


### Grading

Total                                                   | 20 pts
--------------------------------------------------------|--------
Questions 1 and 2                                       | 4 pts * 2 = 8 pts
Questions 3 and 4                                       | 5 pts * 2 = 10 pts
Code style and informatively named code chunks          | 1 pt
Overall organization                                    | 1 pt
