---
title: "The language of models"
author: "Dr. Çetinkaya-Rundel"
date: "October 5, 2017"
output:
  xaringan::moon_reader:
    css: "slides.css"
    logo: img/sta112-hex.png
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

class: center, middle

# Getting started

---

## Getting started

- Any questions from last time?

- Today: Modeling the relationship between variables
    - Focus on *linear* models (but remember there are other types of models too!)

---

## Data: Paris Paintings

```{r message=FALSE}
library(tidyverse) # ggplot2 + dplyr + readr + and some others
library(broom)
```

```{r message=FALSE}
pp <- read_csv("data/paris_paintings.csv", na = c("n/a", "", "NA"))
```

<div class="question">
What does the data/ mean in the code above? Hint: Where is the data file (the csv) located?
</div>

---

## Prices

<div class="question">
Describe the distribution of prices of paintings.
</div>

```{r, fig.height=4, fig.width=7}
ggplot(data = pp, aes(x = price)) +
  geom_histogram(binwidth = 1000)
```

---

class: center, middle

# Modeling the relationship between variables

---

## Models as functions

- We can represent relationships between variables using **functions**

- A function is a mathematical concept: the relationship between an output
and one or more inputs. 
    - Plug in the inputs and receive back the output
    - Example: the formula $y = 3x + 7$ is a function with input $x$ and output $y$,
    when $x$ is $5$, the output $y$ is $22$
    ```
    y = 3 * 5 + 7 = 22
    ```

---

## Height as a function of width

```{r warning = FALSE, fig.height=4, fig.width=7}
ggplot(data = pp, aes(x = Width_in, y = Height_in)) +
  geom_point() +
  stat_smooth(method = "lm") # lm for linear model
```

---

## Vocabulary

- **Response variable:** Variable whose behavior or variation you are trying to understand, on the y-axis (dependent variable)

- **Explanatory variables:** Other variables that you want to use to explain the variation in the response, on the x-axis (independent variables)

- **Predicted value:** Output of the function **model function**
    - The model function gives the typical value of the response variable
    *conditioning* on the explanatory variables

- **Residuals:** Show how far each case is from its model value
    - $residual = actual~value - predicted~value$
    - Tells how far above/below the model function each case is

---

## Residuals

<div class="question">
What does a negative residual mean? Which paintings on the plot have have negative 
residuals?
</div>

```{r warning = FALSE, fig.height=4, fig.width=7, echo=FALSE}
ggplot(data = pp, aes(x = Width_in, y = Height_in)) +
  geom_point() +
  stat_smooth(method = "lm")
```

---

## Multiple explanatory variables

<div class="question">
How, if at all, the relatonship between width and height of paintings vary by whether
or not they have any landscape elements?
</div>

.pull-left[
```{r warning = FALSE, fig.height=4, fig.width=6, echo=FALSE}
ggplot(data = pp, aes(x = Width_in, y = Height_in)) +
  geom_point(alpha = 0.4) +
  stat_smooth(method = "lm")
```
]

.pull-right[
```{r warning = FALSE, fig.height=4, fig.width=6, echo=FALSE}
ggplot(data = pp, aes(x = Width_in, y = Height_in, color = factor(landsALL))) +
  geom_point(alpha = 0.4) +
  stat_smooth(method = "lm") +
  labs(color = "landscape")
```
]


---

## Just for reference...

Here is the code for the two plots in the previous slide

```{r warning = FALSE, fig.height=4, fig.width=6, eval=FALSE}
# points not colored by landsALL type
ggplot(data = pp, aes(x = Width_in, y = Height_in)) +
  geom_point(alpha = 0.4) +
  stat_smooth(method = "lm")
```

```{r warning = FALSE, fig.height=4, fig.width=6, eval=FALSE}
# points colored by landsALL type
ggplot(data = pp, aes(x = Width_in, y = Height_in, 
                      color = factor(landsALL))) +
  geom_point(alpha = 0.4) +
  stat_smooth(method = "lm") +
  labs(color = "landscape")
```

---

## What's a `factor`?

Factor is what R calls categorical variabkles. The `factor` function converts a variable to a factor.

```{r}
class(pp$landsALL)
```

```{r}
class(factor(pp$landsALL))
```



---

## Models - upsides and downsides

- Models can sometimes reveal patterns that are not evident in a graph of the
data. This is a great advantage of modeling over simple visual inspection of
data. 

- There is a real risk, however, that a model is imposing structure that is
not really there on the scatter of data, just as people imagine animal shapes in
the stars. A skeptical approach is always warranted.

---

## Variation around the model...

is just as important as the model, if not more!

*Statistics is the explanation of variation in the context of what remains
unexplained.*

- The scatter suggests that there might be other factors that account for large parts 
of painting-to-painting variability, or perhaps just that randomness plays a big role.

- Adding more explanatory variables to a model can sometimes usefully reduce
the size of the scatter around the model. (We'll talk more about this later.)

---

## How do we use models?

1. Explanation: Characterize the relationship between $y$ and $x$ via *slopes* for
numerical explanatory variables or *differences* for categorical explanatory
variables

2. Prediction: Plug in $x$, get the predicted $y$

---
class: center, middle

# Characterizing relationships with models

---

## Relationship between height & width

```{r}
lm(Height_in ~ Width_in, data = pp)
```

$$\widehat{Height_{in}} = 3.62 + 0.78~Width_{in}$$

- **Slope:** For each additional inch the painting is wider, the height is expected
to be higher, on average, by 0.78 inches.

- **Intercept:** Paintings that are 0 inches wide are expected to be 3.62 inches high,
on average.
    - Does this make sense?

---

## Relationship between height & landscape features

```{r}
lm(Height_in ~ factor(landsALL), data = pp)
```

$$\widehat{Height_{in}} = 22.68 - 5.65~landsALL$$


- **Slope:** Paintings that have some landscape features are expected, on average,
to be 5.65 inches shorter than paintings that don't have landscape features.
    - Compares baseline level (`landsALL = 0`) to other level
    (`landsALL = 1`).

- **Intercept:** Paintings that don't have landscape features are expected, on 
average, to be 22.68 inches tall.

---

## Relationship between height and school

```{r}
lm(Height_in ~ school_pntg, data = pp)
```

- When the categorical explanatory variable has many levels, they're encoded to
**dummy variables**.

- Each coefficient describes the expected difference between heights in that 
particular school compared to the baseline level.

---

## Correlation does not imply causation!

Remember this when interpreting model coefficients

---
class: center, middle

# Prediction with models

---

## Predict height from width {.build}

<div class="question">
On average, how tall are paintings that are 60 inches wide?
\[ \widehat{Height_{in}} = 3.62 + 0.78~Width_{in} \]
</div>

```{r}
3.62 + 0.78 * 60
```

"On average, we expect paintings that are 60 inches wide to be 50.42 inches high."

**Warning:** We "expect" this to happen, but there will be some variability. (We'll
learn about measuring the variability around the prediction later.)

---

## Prediction vs. extrapolation

<div class="question">
On average, how tall are paintings that are 400 inches wide?
$$\widehat{Height_{in}} = 3.62 + 0.78~Width_{in}$$
</div>

```{r warning = FALSE, fig.height=4, fig.width=6, echo=FALSE}
ggplot(data = pp, aes(x = Width_in, y = Height_in)) +
  geom_point(alpha = 0.4) +
  stat_smooth(method = "lm", fullrange = TRUE) +
  xlim(0, 400) +
  ylim(0, 320)
```

---

## Watch out for extrapolation!

> "When those blizzards hit the East Coast this winter, it proved to my satisfaction 
that global warming was a fraud. That snow was freezing cold. But in an alarming 
trend, temperatures this spring have risen. Consider this: On February 6th it was 10 
degrees. Today it hit almost 80. At this rate, by August it will be 220 degrees. So 
clearly folks the climate debate rages on."<sup>1</sup>  <br>
Stephen Colbert, April 6th, 2010

.footnote[
[1] OpenIntro Statistics. "Extrapolation is treacherous." OpenIntro Statistics.
]

---
class: center, middle

# Measuring model fit

---

## Measuring the strength of the fit

- The strength of the fit of a linear model is most commonly evaluated using $R^2$.

- It tells us what percent of variability in the response variable is explained by 
the model.

- The remainder of the variability is explained by variables not included in the 
model.

- $R^2$ is sometimes called the coefficient of determination.

---

## Obtaining $R^2$ in R

- Height vs. width
```{r}
m_ht_wt <- lm(Height_in ~ Width_in, data = pp) # fit and save
glance(m_ht_wt)$r.squared # extract R-squared
```

Roughly 68% of the variability in heights of paintings can be explained by their
widths.

- Height vs. lanscape features
```{r}
m_ht_land <- lm(Height_in ~ landsALL, data = pp)
glance(m_ht_land)$r.squared
```

---

## Hands on!

See Slack for brief activity link.
