---
title: "Case study 01 - Model selection for Paris Paintings"
date: "Due at the end of class on Oct 19"
author: "Team assignment"
output: 
  html_document: 
    theme: yeti
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Accessing the data

The data can be found in the `\data` folder of your homework repository.

Read the data with

```{r eval=FALSE}
pp <- read_csv("data/paris_paintings.csv")
```

Take a peek at the [codebook](https://www2.stat.duke.edu/courses/Fall15/sta112.01/data/paris_paintings.html).

### Accessing the assignment repo

Go to the #assignment-links channel on Slack and click on the link for mini-hw-07, and accept the assignment.

### Task

In this case study your goal is to fit models that include both main and interaction effects and do model selection. 

#### Planning:

1. Decide on a subset of variables to consider for your analysis. Think 
about it as focusing on certain aspects of the price determination, as opposed
to all aspects. You're allowed a maximum of 10 total variables to consider, 
including interactions. The more variables you consider the longer model 
selection will take so keep that in mind.

2. Decide among these which variables might make sense to interact. Remember, 
we consider interactions when variables might behave differently for various 
levels of another variable. Ideally, you would get some expert guidance for
choosing which variables to interact, however the curators of the data are not 
in class with you, so instead they have compiled a list of interactions that 
might be potentially interesting:

    - School of painting & landscape variables: `school_pntg` & `landsALL` / `lands_figs` /
    `lands_ment`
    - Landscapes & paired paintings: `landsALL` / `lands_figs` / `lands_ment` & `paired`
    - Other artists & paired paintings: `othartist` & `paired`
    - Size & paired paintings: `surface` & `paired`
    - Size & figures: `surface` & `figures` / `nfigures`
    - Dealer & previous owner: `dealer` & `prevcoll`
    - Winning bidder & prevcoll: `endbuyer` & `prevcoll`
    - Winning bidder & artist living: `winningbiddertype` & `artistliving`

This is not an exhaustive list, so you might come up with others.

#### Model fitting and selection:

3. Use backwards elimination to do model selection. Make sure to show 
each step of decision (though you don't have to interpret the models at 
each stage).

#### Model interpretation:  

4. Provide interpretations for the slopes for the final model you arrive at 
and create at least one visualization that supports your narrative.

### Grading

Check / no check