---
title: "HW 03 - Simulating sampling distributions and null distributions"
date: "Due: Nov 14 at 10:05am"
output: 
  html_document: 
    theme: yeti
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Tasks

<!-- Part 1: Sampling distributions of sample means -->

### Part 1: Sampling distributions of sample means

1. Load the following dataset:

```{r read-num-pops}
num_pops <- read_csv("data/num_pops.csv")
```

2. Each column represents a different population. Determine out how many columns 
are in this dataset. Create a histogam of each of the populations, and describe
their distributions. Specifically make sure to note the shape, center, and 
spread of the distribution.

3. Create a sampling distributions of sample means using *samples* of the 
following sizes for each of the populations:

    - n = 10
    - n = 50
    - n = 200

Generate each sampling distribution using at least 15,000 samples. Note that sampling distributions are created by taking random samples, with replacement, from the original population.

To help you out, here is the code for creating a sampling distribution for samples
of size 5 with 20 samples for the population in column M of the data. Also make sure 
to read the help for the `rep_sample_n` function. Note that this function is 
in the `oilabs` package.

```{r mean-sampling}
sampling_mean_n5_rep20_M <- num_pops %>%
  select(M) %>%                                          # select column M
  rep_sample_n(size = 5, reps = 20, replace =  TRUE) %>% # 20 random samples w/ replacement,
                                                         # each of size 5
  group_by(replicate) %>%                                # group by each replicate
  summarise(xbar = mean(M))                              # calculate the mean of each sample
```

Make sure to use clear naming convention for your sampling distributions. You can 
follow the naming convention I used above for `sampling_mean_n5_rep20_M`, or choose a 
different, but reasonable, style.

3. For each sample size create two graphs:
    - A visualization of a single sample.
    - A visualization of the sampling distibution.

4. Describe the shapes of your sampling distributions; calculate the centers (mean) 
and the spreads (standard deviation). Compare these to shapes, centers, and spreads 
of the parent population distributions from (1).

5. Note the Central Limit Theorem for a single mean For each one of your samples,
evaluate whether the Central Limit Theorem holds. If it does not, explain why.

<!-- Part 2: Sampling distributions of sample proportions -->

### Part 2: Sampling distributions of sample proportions

1. Load the following dataset:

```{r read-cat-pops}
cat_pops <- read_csv("data/cat_pops.csv")
```

2. Each column represents a different population. Determine out how many columns 
are in this dataset. Create a bar plot of each of the populations, and describe
their distributions. Specifically make sure to note the proportion of successes.

3. Create a sampling distributions of sample proportions using *samples* of the 
following sizes for each of the populations:

    - n = 10
    - n = 50
    - n = 200

Generate each sampling distribution using at least 15,000 samples. Note that sampling distributions are created by taking random samples, with replacement, from the original population.

To help you out, here is the code for creating a sampling distribution for samples
of size 5 with 20 samples for the population in column A of the data.

```{r prop-sampling}
sampling_prop_n5_rep20_A <- cat_pops %>%
  select(A) %>%                                          # select column A
  rep_sample_n(size = 5, reps = 20, replace =  TRUE) %>% # 20 random samples w/ replacement,
                                                         # each of size 5
  group_by(replicate) %>%                                # group by each replicate
  summarise(phat = sum(A == "success") / n())            # calculate the mean of each sample
```

Make sure to use clear naming convention for your sampling distributions. You can 
follow the naming convention I used above for `sampling_prop_n5_rep20_A`, or choose a 
different, but reasonable, style.

3. For each sample size create two graphs:
    - A visualization of a single sample.
    - A visualization of the sampling distibution.

4. Describe the shapes of your sampling distributions; calculate the centers (mean) 
and the spreads (standard deviation). Compare the centers of your sampling distributions 
to the true population proportions from Question (1)

5. Note the Central Limit Theorem for a single proportion. For each one of your samples,
evaluate whether the Central Limit Theorem holds. If it does not, explain why.

<!-- Part 3: Simulations for hypothesis testing -->

### Part 3: Simulations for hypothesis testing

Describe precisely how you would set up the simulation for the following hypothesis tests. 
Imagine using index cards or chips to represent the data. Also specify whether the null 
hypothesis would be independence or point and whether the simulation type would be 
bootstrap, simulate, or permute. In each of the scenarios you can assume sample size is
100 and number of simulations is 15,000.

- Describe the simulation process for testing for a single population standard deviation. 
Suppose the research question is asking whether the standard deviation is less than 3, 
and the observed sample proportion is 2.
- Describe the simulation process for testing for the difference between two population proportions. Suppose the sample proportion of successes for sample A is 0.3 and for 
sample B is 0.4.
- Describe the simulation process for testing for a single population proportion. Suppose 
the research question is asking whether proportion of successes is majority, and also that 
the observed proportion of success is 0.6.
- Describe the simulation process for testing for the difference between two population means. Suppose sample mean for sample A is 5 and for sample B is 3.

## Tips

- You're working in the same repo as your teammates now, so merge conflics will happen, issues will arise, and that’s fine! Commit and push often, and ask questions when stuck.

- Review the grading guidelines below and ask questions if any of the expectations are unclear.

- Make sure each team member is contributing, both in terms of quality and quantity of contribution (we will be reviewing commits from different team members).

- Set aside time to work together and apart (physically).

- When you're done, review the .md document on GitHub to make sure you're happy with the final state of your work. Then go get some rest!

## Grading

Total                                                   | 100 pts
--------------------------------------------------------|--------
Part 1: Sampling distributions of sample means          | 30 pts
Part 2: Sampling distributions of sample proportions    | 30 pts
Part 3: Simulations for hypothesis testing              | 20 pts
Code quality                                            | 10 pts
Document organization  (team name, code chunk names, commtis, overall organization, etc.)                                     | 10 pt
